# A. Please start here {.unnumbered}

```{r message=FALSE, warning=FALSE}
#| echo: false

source(here::here("R/functions.R"))
```

## Preface

I began drafting some distance-learning material based on Dr Deming’s invaluable teaching as long ago as 1999. But various events, combined with lack of time, interrupted that work and, when I retired at the end of 2004, it had remained untouched for almost five years.

These past few years have been very different. With the encouragement of my old friend Richard Capper, who some 25 years ago founded a superb little consultancy bravely named The Universal Improvement Company (UIC), I revisited those early jottings in 2012. Richard had always sent recruits to his staff to my public seminars, but these had, of course, now been unavailable for over seven years. So, when he contacted me in January 2012, Richard first tentatively explored whether I could come sufficiently out of retirement to present some in-house seminars for him and his people: but I knew I had grown far too rusty to attempt that! Instead, I offered to try to tie up some loose ends in the work that I’d left in 2000 and hand it over to him to use with his staff and clients as he saw fit. He enthusiastically accepted the offer.

However, soon after I got to work, I began to wonder how many others were in a similar position to Richard—keen to introduce their colleagues to Deming’s work but regrettably feeling they had no good means of doing so. Also, there must be many others who come across Deming, perhaps in some books that they are reading about management and quality, etc, or by finding some videos of him while browsing the internet. But, if they get interested, how can they really learn something substantial—in both breadth and depth—about what he taught? Just from reading a few books or watching a few videos? I don’t think so; I know that wouldn’t have worked with me. But, extremely fortunately, I had the relatively rare opportunity to attend and be involved with many of Dr Deming’s four-day seminars. And even then, to be honest with you, it was only at the third of those seminars that his teaching really began to take some shape in my mind. Before then I had remained pretty puzzled and confused—it was so different.

Incidentally, that’s as good a reason as any about why I have laid out this course in the form of “12 Days” of learning. If it took 3 × 4 = 12 days of Dr Deming’s own teaching for his wisdom to begin to get through to me, how could I expect anyone else to need anything less?

A few months after that third four-day seminar, I began to develop my own teaching on Dr Deming’s work—tentatively at first to my students at the University of Nottingham and then, before long, offering some short introductory public seminars. Subsequently, during the nearly 20 years before I retired, I was privileged to present many hundreds of seminars and courses—thus involving many thousands of delegates and students—that were solidly focused on Dr Deming’s teaching.

By the end of 2012, I had begun to feel confident that I could incorporate much more of the experience that I had thus gained over those years into what I had started back in 1999. But I also realised I would need a lot of help. I thus contacted a number of old friends and colleagues to let them know what I was up to and to ask if any of them might be interested in reading what I was drafting and hopefully giving me some feedback and guidance on how I could improve it. It was one of the best decisions in my life! The amount of constructive feedback, along with extremely useful information and the many contributions to the course that I have received during the subsequent years, has transformed those early jottings into some teaching and learning material that has been greeted with enthusiasm and excitement both by those who have been helping me and others who have seen the way the material has been developing during the past few years.

I would also like to thank those connected with the UIC who had not previously seen any of the material but bravely volunteered in 2017 to become my initial “guinea-pig” students to try working through the complete course. Also, regarding the UIC, my special gratitude goes to my main contact there since early 2016, Lucy O’Melia. Lucy aided the progress of the course in innumerable ways, including exploring matters of design and beginning the hefty task of seeking approvals for my use of materials from a host of other sources during the course. This task turned out to be extremely time-consuming and was then continued and successfully concluded by Luca Willington of the UK’s Deming Alliance. I am indebted to both these ladies for the considerable time and effort they have contributed to bringing 12 Days to Deming to this current state of development which I am now able to make available to everybody interested.

So it is my great pleasure to acknowledge those who have given freely of their time, several of them ever since 2013, to aid me in this project in so many and various ways. I’m certainly not going to try to tell you individually how every one of them has helped and contributed—otherwise this Preface would be at least twice as long as it is now. Here therefore, simply in alphabetical order, is a list of those who have been particularly involved, to all of whom I owe my deep and profound gratitude: Mitch Beedie, Richard Capper, the late Nigel Clements, the late John Dowd, Eric Lawson, Andrew Forrest, Malcolm Gall, Nick Gardener, Jackie Graham, Dave Higgs, Scott Hindle, Alan Hodges, Dave Kerr, Sarah Kerridge, Karl Laidman, Brian Leeming, Arvind Mathur, Tony Miller, Lucy O’Melia, Balaji Reddie, the late Peter Scholtes, Ray Tew, Richard Thorns, Lou Tribus, Fran and Don Wheeler, Fiona Wilkie, Colette Williams, Luca Willington, David Wormald, Peter Worthington and Dave Young. I fear I must still have omitted the names of others who should also be included in this list: please put it down to absent-mindedness due to advancing years.

There is another lady to whom I am also inordinately grateful: my better half, Tonie. During these past seven or eight years, she has, of course, put up with such considerable lack of time and attention from me that most certainly could not have been anticipated or wanted over so many years of what was supposed to be my retirement. But, despite that, she’s still here with me—bless her!

And finally, of course, to Dr W Edwards Deming. Without doubt, I was a slow learner. But the patience and respect which he showed toward me throughout those eight and a half years during which I was so fortunate to know him and learn from him is still, frankly, inexplicable to me. I treasure many memories of those happy times.

On one occasion that I met up with him while I was writing my book: *The Deming Dimension*, I asked him whether he minded my quoting him as extensively and explicitly as I was beginning to do. His answer was the greatest compliment that I could ever have imagined. He simply replied that I would always have his permission to use any of his material, written or spoken, in whatever way I wished. Well, I have made considerable use of his exact words in 12 Days to Deming, particularly in the Second Project which occupies the whole of Days 10 and 11—the climax of the learning in this course. I hope so much that the context in which I have used his words there and throughout the course justifies the extraordinary confidence and trust that he showed in me. In turn, it is my earnest hope that students who embark upon this course will grow to value and gain from that same wisdom as much as I shall always do.

Henry Neave, January 2020

## Active learning

Why an “active-learning” course? You are probably more familiar with the term “distance-learning.” However, distance cannot improve learning, but activity certainly can.

There is, of course, activity in normal taught courses, be they provided at school, college, university, or any other type of educational institution. The activity may be homework, “prep,” essays, assignments, perhaps projects or even minor research. I think most would agree that, although the lessons, lectures, background reading, etc. are essential, the learning you gain from them often does not become particularly deep-seated in your mind until you work at those subsequent activities, using and applying that initial learning. “Doing something” with the initial learning helps it to become real.

Adding to your learning by reading a book about a topic with which you are already somewhat familiar is often absolutely fine: you already have the foundations upon which to build. But trying to learn something “from scratch” just by reading a book—however good the book is and whoever wrote it—quite often does not work as well. I have designed this course on the presumption that you will indeed be working from scratch, i.e., that you have literally “no previous knowledge” of anything specifically related to the subject matter. It is up to me to help you learn what I’d like you to know—not expect you to know it already!

In some areas, e.g., Mathematics, Engineering, and some Sciences, books may include questions and exercises at the end of each chapter, and they can help. But, as with homework, etc., such activity often takes place some time after the initial exposure to whatever is being studied, not during it—i.e., not as an integrated component of the learning process. It is when the activities do take place as a fully integral part of the learning process that I use the term “active learning.” And that’s quite different from weekly assignments or exercises at the end of a chapter: in active learning, the activities are carried out as and when they are most appropriate and helpful, not hours or days or even weeks or months later.

The amount and extent of such activities can be highly variable depending upon the context. For instance, when covering relatively gentle background material, hardly any such activity may be needed. At the other extreme, there can be situations where the learning is almost entirely activity-based. So there will be some occasions during this course when the activities are nearly continuous: you will be reading material maybe even just a sentence at a time and then immediately “doing something” as a result of that sentence. On other occasions, you will simply be reading for quite a while. As you’d expect, most of the time, you’ll be somewhere between those two extremes.

## Time management

### Days

It is because of this wide variety in the forms of learning that I have laid out this course in the form of 12 “days” rather than 12 chapters or 12 modules, etc. (I appreciate that it would be the rare student indeed for whom the “12 days” will be twelve consecutive days!) By a “day” I am envisaging a full working day of
up to eight hours, though sometimes a little less. Of course, you individually or the group with whom you are working can adapt this timeframe as suits you. But if I were teaching you this material in person, naturally leaving suitable amounts of time for you to carry out the various activities, then this is the kind of timeframe that would seem reasonable to me.

I also appreciate that it might not be possible or convenient for you to devote full days to this learning. For instance, if you are quite busy with other things then half-days are likely to be more suitable for you. Nevertheless, because of this context of “days”, I do often refer to “morning” or “afternoon”, etc—but you’ll
easily be able to interpret such references in ways that are relevant to your particular schedule. Indeed, a “half-day” could well be an evening rather than a morning or afternoon—in which case I’d suggest your evening should start quite early unless you are a real “night-owl”!

If at all possible, I recommend you do not attempt shorter sessions than half-days. This material is very much a case of “joined-up thinking”: it would not take kindly to being split up into lots of little bits and pieces. I would hate to try to teach it in, say, separate lessons or lectures typically of an hour or less. Fortunately, I never had to try! My seminars, both public and in-house, were always scheduled over one or more days, and even my University courses on Dr Deming’s work were presented as four-day seminars spread over two weekends.

Obviously, what you decide upon will depend on whether you (and your colleagues if you are studying in a group) are already involved in other full- or part-time study or other kind of work. If you are in a group then naturally a schedule of regular meetings will need to be organised. But, even if you are studying this course on your own, you would be wise to consider upfront the kind of time you will have available and then come up with a regular schedule that you’ll be able to maintain. Many people who are in full-time work but who simultaneously take on a course of study, e.g. for some additional qualifications (like many Open University students in the UK), are rather used to doing their full-time work during the week and then studying at the weekends; so maybe for you to set aside Saturdays or Sundays for this course might be more practical and appealing. On the other hand, if currently you do not have any full-time commitments and therefore could consider taking on this course more intensively, you might consider two or even three “days” per week. But the latter would indeed be quite intensive! On the other hand, because of the “joined-up” nature of the material, I recommend that you should aim to cover no less than one day’s work per week if at all possible—i.e. covering the course over a three-month period—otherwise you’ll be in danger of losing the continuity and momentum which will help this learning to succeed. An obvious good compromise between
those two possibilities would be six two-day weeks.

### Too much time, too little time

Let’s move on now to considering how to manage the time within each of our “days”. Again you can be quite flexible as suits you personally or the group with whom you are working. But I would advise that spending either considerably too little time or even too much time on anything here is likely to be unwise.

If you spend too little time then obviously you are not giving the learning sufficient opportunity to “sink in”. Spending too little time is especially dangerous with Deming’s work, particularly when you are reading his own words—which will be happening quite a lot during the later stages of the course. He was brilliant at saying a great deal in very few words—but that has its pros and cons. This is not material for the speed-reader. This is material for the student who has the patience to ponder and dig deep: for there is plenty to uncover!

I particularly recommend that you do not try to rush through the work on Days 4 and 5. This course contains two substantial projects. The First Project is designed to keep you busy throughout the afternoon of Day 4 and all of Day 5. Apart from being very important in its own right, essentially it also serves as a training-ground for the deepest-thinking part of the whole course: namely, the Second Project which takes
up almost the whole of Days 10 and 11. Good work in the First Project will prove to be extremely helpful preparation for the Second Project.

But again, don’t go to extremes. Rather less obviously, spending considerably too much time, conscientious as that would show you to be, also has its drawbacks. Again you would then be in danger of losing
the continuity and momentum that I’ve already mentioned. I know that, almost certainly, there will be occasions when you reach something of a “mental block” and e.g. find you cannot make progress with a particular activity even after spending some reasonable amount of time on it. But, extensive and varied as
Deming’s teaching is, you will increasingly discover that it does all fit together. That is indeed one of its
tremendous strengths—it is not just some miscellaneous concoction of isolated ideas. So it is quite likely
that, if you set aside that troublesome matter for the time being and move on to whatever follows, you may
soon come across something that will “make the penny drop” regarding what you were puzzling over. So
then you can go back and really get moving on that troublesome activity. Furthermore, as the course progresses, several of the activities will become increasingly open-ended. In some such cases it might be
feasible for you either to just jot down a few quick notes or, at the other extreme, maybe develop an article
for publication on the topic! But, while this course is in progress, I’d prefer you to aim at somewhere in
between those extremes!

One consequence of the “joined-up” nature of Deming’s work is that you will see literally hundreds of page
references during this material, references both to elsewhere in the course and to other sources. You are
not expected to follow up all of those references! They are there for your guidance and help, either during
the course or later, simply as and when you feel you might find them most useful. I should also point out
that the pages for each day, and in other sections such as this Welcome and the Appendix, are all numbered from page 1 upward rather than being continuous throughout the course. The reason is that, very
early on while developing this material, I discovered that this made it much easier for me to find my way
around—and I hope the same will prove to be true for you.

### Detailed guidance

I’ll begin here with a great big caveat: the following guidance is included for you very much on a “take it or
leave it” basis. If you are experienced with distance-learning courses or other types of largely unsupervised study then you may well decide to “leave it”—and that’s fine by me since in that case you will already
have become used to managing your own time and will know the speed and style of learning that works for
you. So then you may feel that the kind of guidance I’ll describe here is too restrictive for you or maybe
even patronising! No problem. As with all those page references just mentioned, please feel free to ignore
my guidance unless you feel it will be helpful to you. On the other hand, if your past experience has been
wholly or largely in the form of taught courses in schools and colleges etc, it can be quite tricky to handle a
self-study course that leaves you on your own without the kind of supervision and timetabling which naturally accompanies taught courses. So, if that is your situation, I believe you may well find my guidance to be
useful.

As many people in my position would do, I’d invariably include some personal guidance about timings
when writing up my own notes in preparation for delivering courses, seminars or conference presentations. 
The timing for any particular section would not necessarily be reflected in the length of my notes for that
section: some material can be dealt with quite quickly and easily while other material needs to be covered
with greater deliberation. I found such indications of timing to be extremely useful—as long as I did indeed
treat them just as approximate guidance rather than as anything more regimented! I have thus imagined
myself presenting this course to you in person and similarly including some timing indications for my own
guidance—except that now, of course, in practice that guidance is for you if you’d like it. In particular, if
presenting the course in person I would need to allocate (following what I have just said) neither too little
nor too much time for the activities, and I’m hoping you will find my guidance on timings to be particularly
helpful with them. Also, returning to the reading-matter, you may often find yourself moving through it
rather faster than I am indicating. But just zooming through it is not the object of the exercise: sufficient
time is needed for it to penetrate! That may require reading some of the material relatively slowly and carefully, perhaps re-reading some of it, and maybe writing up a few notes on it. In my guidance I’ll try to leave
you sufficient time for all of these as appropriate.

As mentioned earlier, I have presumed your available time is up to eight hours per “day”, although with
some days being a little shorter than that. Broadly speaking, the length of the days will tend to increase as
we work through the course (although there will be exceptions). In particular, I am estimating that Day 1
could take only about six hours whereas during the Second Project on Days 10 and 11 there is little doubt
that the full eight hours per day will be needed to even begin to do it justice.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| column: margin

create_clock(6, 00)

create_clock(9, 30)
```


To be specific, I have imagined that we shall usually start at 9.00 am, will take an hour off for lunch, and will
close at some time between 4.30 pm and 6.00 pm. (No coffee or tea breaks—I shall assume that liquid
refreshment is on tap all day long!) A lunch-break is indicated in both the text and in the list of contents for
each day by a red, white and blue box that you will not be able to miss! Little clock icons such as those I’m
showing you against this paragraph will appear both on the opening page for each day against a description of the day’s contents and also at various places within the text. Again, treat these indications as rough
guides rather than instructions. I believe they will help you to pace yourself appropriately through the reading and particularly when carrying out the activities. As indicated above, Day 1 is a little different regarding
the overall length of the day, since there is much less in the way of activities on this opening day of the
course than subsequently. So, in that initial case, I am permitting a later start!

It is, of course, likely that your preferred starting-time will generally, if not always, differ from mine. So, if
you want to keep an eye on my guidance about timing but also prefer to avoid having to keep carrying out
simple arithmetic on the time-difference, it might be sensible to adjust a clock or watch in order to synchronise with my starting-time.

## Self-study or group-learning

Finally, while developing this course, I have primarily been thinking of it as being used for self-study so that,
in particular, the activities are reasonable and feasible for an individual student to carry out. (I have special
empathy for an individual student studying Dr Deming’s work alone because, back in 1985, I was one!)
However, the course is just as suitable for study in small or medium-sized groups meeting together either
physically or via the internet. (There are a few places in the material where I am clearly focusing on the
individual student, but they should not cause members of study groups any difficulty.) An advantage of
being an individual student rather than in a group is that you will, of course, be able to adjust the length of a
“day” to one that suits you—but not forgetting what I’ve said about both too little or too much time.

By “small” groups I am thinking of perhaps just two or three people, maybe four, who are reading through
and discussing the material and carrying out all the activities together. Such small groups (again meeting
physically or virtually) working together cooperatively throughout the entire course can be extremely successful and rewarding. Indeed the very concept of cooperation, working together, win–win is one of the
many crucial themes in Dr Deming’s teaching. But, of course, people differ from one another. For example, some are naturally quite fast learners, some are not. Being in the latter category is not necessarily bad:
slower learners may delve deeper into the learning compared with fast learners who might tend to skate
over the surface. So you can see why I emphasised “cooperation” near the start of this paragraph!

The other option for small groups is the same as I would recommend for medium-sized groups—say up to
ten or twelve people—and that is to enjoy the best of both worlds with a mixture of self-study followed by
meetings for discussion. (With larger groups it becomes increasingly difficult for everybody to participate in
the discussions.) At the time of writing (July 2021) I have recently had the great pleasure of watching and
listening to such an arrangement in action since the organiser of a study-group of a dozen students kindly
videoed all the meetings and sent them to me. It worked fantastically well, despite having what I considered to be a very ambitious schedule of two meetings per week each lasting about 90 minutes and covering either a complete day of the 12 Days or, especially in the second half of the course, a half-day.

## The "Overture"

There is no need for me to summarise the content of the course here since that is provided on Day 1 pages
16–19. However, the morning of Day 1 covers quite a wide range of topics, and so I will now give you a
brief description of what you will find there. Incidentally, if you were wondering about my choice of title for
this first half-day, you will find my explanation upfront on Day 1 page 1!

**Deming (re)discovered** (Day 1 pages 2–4)

This section contains a brief introduction to particular parts of the “Deming story” that are covered more
fully during the afternoon. There is also a mention of Dr Deming’s celebrated four-day seminars which he
presented on literally hundreds of occasions after being “rediscovered” in the West around 1980. The section ends with an indication of some useful Deming-interest contacts in the UK and elsewhere and also on
the internet.

**"Statistics? Oh no!"** (Day 1 pages 5–10)

The subject of Statistics is hardly everybody’s favourite area of study! But the truth is that, in essence, Dr
Deming’s life’s work was launched as the result of his learning from Dr Walter A Shewhart in the 1920s
about a new, practical and illuminating approach to interpreting process-data. (By “process-data” I simply
mean any data that are recorded over time at fairly regular intervals—daily, hourly, weekly, every minute,
etc). That being the case, nobody can get very far into understanding Dr Deming’s work without knowing
something about what Shewhart came up with.

But, fear not—this does not have to involve you in any heavy Mathematics! In fact, during this course,
computational work is only involved on Days 2 and 3—so those two Days are the only occasions during the
course when you may need to use a calculator! But, even then, you are free to cut down such arithmetic to
a bare minimum if you so wish: you will not be at any disadvantage as regards understanding the main
content of this course if that is what you choose. The disadvantage would only arise when you subsequently try to put certain parts of the learning from the course into practice, as at that stage it would be
very handy if you had learned how to study and interpret process-data in the way that Shewhart developed.
The reason is that an essential component of putting what Deming taught into practice is the ability to
improve processes (of whichever kinds are relevant in your work—administrative, service, production, management, financial, training, etc). And, of course, you will probably need to study and interpret some data
from a process in order to help you to improve it.

The delegates at my seminars came from all sorts of different backgrounds and experience: they ranged
from the many who would prefer to run a mile away from anything “technical” to a group from the British
Government’s Central Statistical Office! So I got used to catering for such a wide range of interests and
abilities, and have also done so in this course. This section starting on Day 1 page 5 tells you how.

**Helpful books and videos** (Day 1 pages 11–13)

That title speaks for itself. The only essential additional source for the course is the prescribed text: my
book The Deming Dimension which was first published in 1990—for some information about it please see
pages 12–13 later in this Welcome section. However, as with pretty much any course on any subject, there
are numerous other sources that can be helpful, and this section suggests and briefly describes a few of
them.

**Activities, Major Activities, Pauses for Thought, Projects and "Your Organisation"** (Day 1 pages 14–15)

This section describes the various types of activities that make this course an active-learning course. Several of these, especially in the second half of the course, encourage you to relate what you are learning to
what is happening in whatever company or other type of organisation you are currently working in—and
also suggests what to do if you are not currently involved with any organisation.

**Outline of the course** (Day 1 pages 16–19)

This outline has already been mentioned Immediately beneath The "Overture" on the previous page.

**Quality, theory, philosophy, management...** (Day 1 pages 20–22)

This section introduces you to some of the terms and expressions that will occur quite frequently during the
course and also provides you with your first three “Pauses for Thought”.

**The attraction of Deming’s work** (Day 1 pages 23–26)

So what’s so special about what Dr Deming taught? This and the following section should give you some
ideas about that.

**Deming is different** (Day 1 pages 27–28)

And how! If perhaps you thought that what he taught was just a little bit different from what you may have
learned elsewhere about quality and management, etc, this section might soon change your mind!

**Some light relief!** (Day 1 page 29)

Just in case you regard some of the previous section’s content to be a little crazy, here you’ll see some
other illustrations which were originally regarded as crazy but turned out not to be so crazy after all.


## Serving both the "Technophobes" *and* the "Technophiles"—


