---
title: "'STATISTICS? OH NO!'"
execute:
  echo: false
---
  
```{r}
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)

source(here::here("R/functions/main-functions.R"))
```

# Statistics?

Now, depending on your background, one feature of the previous section may have already started making
you feel a little queasy. And that is the repeated references to things *statistical*. You know that Dr Deming
described at least some of his work as “statistical studies”. You know that one of his first recommendations to Bill Conway was that he hire an experienced statistician. You know that the “all-important” control
chart is a statistical tool. And, perhaps worst of all, you know that I used to teach Mathematical Statistics!

I can heap yet more coals onto the fire by quoting the way that Dr Deming was introduced midway through
*If Japan Can, Why Can’t We?*. These are the words of the programme’s
narrator, Lloyd Dobyns:

>We have said several times that much of what the Japanese are doing [is what] we taught them to
do. And the man who did most of the teaching is W Edwards Deming, statistical analyst, for whom
Japan’s highest industrial award for quality and productivity is named. But in this country he is not
widely recognised. That may be changing.

“W Edwards Deming, statistical analyst”? Well, if so, be sure that that was but a single string on a many-
stringed bow!

<span class=neave_note>As you may already have noticed, I am using British spellings and punctuation conventions even when
quoting Americans. Since, obviously, I quote Americans a great deal in this course, I thought it appeared
rather pedantic and messy to keep switching between the two styles. For simplicity I therefore decided to
use British conventions throughout. I hope this causes no offence or difficulty.</span>

But no wonder then that, if you do not have any background in Statistics, you may be beginning to feel a
trifle uneasy. But even if you *have* suffered some introductory Statistics course either recently or long ago,
you might also be feeling uneasy. For I am well aware that the subject of Statistics is not a favourite area of
study for many people! And yes, I know all the old sayings such as “There are lies, damned lies, and statistics” and “You can prove anything with statistics”, and I am aware of (indeed, have read and enjoyed)
books such as Darrell Huff’s best-seller: *How to Lie with Statistics*.

However, *all* we see in the main material of this course that is related to the subject of Statistics is what
I like to summarise in just two words: “understanding variation”. In fact, those two words form the short
title of the best introductory book on the topic that I know of. The full title of that book is also instructive:
*Understanding Variation—the Key to Managing Chaos* (written by my good friend Dr D J Wheeler). But no,
I’m not expecting you to become expert in the subject-area known as “Chaos Theory”! In a nutshell,
understanding variation is to do with being able to justifiably describe the behaviour over time of processes
or systems of any kind by words such as “stable” and “predictable” or, on the other hand, “unstable” and
“unpredictable”. Those words indicate pretty well the difference between the two states, and I think you
will easily understand that, if your work involves what we are referring to as “processes” or “systems” (and
most work certainly does) then that difference is pretty darned important! And the control chart is the
invaluable tool which best enables us to discriminate between those two states. The “official” terms for the
two states that you will find both Drs Shewhart and Deming using are respectively “in statistical control”
and “out of statistical control” although, in the former case, Deming also often refers to a “stable system”.

The point I want to emphasise here is that what I am expressing as “understanding variation” is *different*
from the material usually taught in introductory (and, indeed, later) conventional Statistics courses. If you
have ever had an introductory course in Statistics then I doubt very much whether you ever heard of either Walter Shewhart or the control chart during that course. Indeed, it is sadly true that some people who have
university degrees in Statistics haven’t heard of them either! In this course (apart from the Optional Extras
section which is included purely for the benefit of people who might be interested in more technical matters) we shall *not* be involved with such things as probabilities or the normal distribution or the binomial distribution or any other probability distribution, nor with tests of significance, *p*-values or confidence intervals,
nor with techniques to do with regression and correlation, etc, etc. We do not need them. The large majority of what is normally taught in traditional Statistics courses is irrelevant for “understanding variation” in
the all-important sense in which we use the term in this course. *So, if you haven’t heard of the list of things
I’ve just detailed, you actually have the advantage: for then you don’t have any **un**learning to do.*

<span class=neave_note>That comment might sound flippant. It isn’t. Over the nearly 20 years of my seminars on Dr Deming’s
teaching I rarely suffered from having any “difficult” delegates. The few that I had could be divided into
two types. One type were very senior managers, the other type were those with some qualification in
Statistics. Of course, I don’t mean *all* senior managers or all statisticians: I mean those who came to the
seminar with the impression that they already knew all they needed to know—so really there wasn’t much
point in their attending! Regarding statisticians, note well my first quotation below.</span>

But, with my background as a university lecturer in Mathematical Statistics, I didn’t realise any of the above
when travelling to London in June 1985 for that first four-day seminar with Dr Deming. You might be
amused by a paragraph I wrote on page 76 of the second edition of my little book of *Statistics Tables* (published in 2011):

>At that first seminar, I was looking forward to discovering how all my knowledge of Mathematical
Statistics fitted into Dr Deming’s management teaching. It didn’t! I believe there was not a single
mention of a probability, a probability distribution or a hypothesis test during the whole four days.
Yet Deming was a *statistician*?! (And indeed a fine mathematician as well.) Further, the *only* statistical technique he used during the whole four days was the control chart … I had new and very different learning to face.

Incidentally, the full title of *Statistics Tables is Statistics Tables for Mathematicians, Engineers, Economists
and the Behavioural and Management Sciences* which, for obvious reasons, I shall henceforth abbreviate
by *ST*! And, while I’m about it, I may as well also quote a couple of paragraphs from the “Welcome to the
Second Edition” on page 2 of my *Elementary Statistics Tables* (to be similarly abbreviated by *EST*), which
was also published in 2011:

>So, what’s new? The majority of the added material focuses on a remarkable statistical technique
which, to all intents and purposes, was unknown at the time of the original edition. At this time of
writing it is still largely unknown, especially in academia—it hasn’t yet reached most of the introductory Statistics books and courses. But, during the second half of my career (mostly spent outside
academia, unlike the first half), I found the *process behaviour chart* <span style="font-family: serif; color: #333; font-size: 16px">[see the small print below]</span> unbelievably useful due to its (I believe) unique combination of simplicity and effectiveness. Of course,
you won’t find it on examination papers. But if you want to analyse and understand data out there
in the ‘real world’, I believe you’ll find it invaluable. Many delegates, sent by their boss, would arrive
at my public and in-house seminars in fear and trepidation: they’d never been able to ‘do Statistics’—they hated the subject! By the end of the day they could understand the process behaviour
chart, and they could use it, and they could *communicate* with it.
So here it is in this new edition of *Elementary Statistics Tables*—though you don’t even need
any tables in order to use it! Don’t scorn its simplicity—try it out on some real process data, including some that you have not previously attempted to interpret.

:::: {.columns}

::: {.column width="85%"}
<span class=neave_note>A couple of points of clarification will be useful here. Firstly, the original editions of these little books of
Statistics Tables were published as long ago as 1978 and 1981 respectively—long before I met Dr Deming or learned anything significant about his work. Not surprisingly therefore, those original editions contained little on control charts because, back then, I honestly had no idea of how important control charts
are in practice. Second, I’ll remind you of something I mentioned on page 7 of the Welcome section:
“process behaviour chart” is an alternative term for “control chart” which nowadays is preferred by many
people who use the chart for the purposes relevant to Dr Deming’s teaching. Indeed, I also have generally used it in recent years because, compared with usual interpretations of the word “control”, it describes much better why the chart is so useful. The reason I’m calling it a control chart in this course is
that I shall be including many direct quotations from both Drs Deming and Shewhart. Naturally, they used
the traditional language, and so, similarly to something that I said on page 5, it didn’t seem sensible here to
keep flitting between alternative terms for the same thing.</span>
:::

::: {.column width="15%"}
<div style="margin-top: 290px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(9, 55)
```
</div>

:::

::::

# The two paradoxes

So, as implied by that quote from *EST* and earlier, the truth is that the statistical tool which had pride of
place in Dr Deming’s teaching is rarely to be seen in introductory courses on Statistics—nor, for that matter,
in more advanced Statistics courses. I can go further: *the bulk of material which is featured in such courses
does not get mentioned in Dr Deming’s teaching—except, in some cases, for him to point out that it is
unnecessary and misleading!* Quite some paradox. And it’s not the only one.

How you react to that first paradox is likely to largely depend on whether or not you have any background
in what I’m calling “conventional” or “traditional” Statistics. If you haven’t then I imagine you will welcome
the paradox with open arms! But otherwise you may have more of a sense of puzzlement rather than relief.
If that is the case then I can sympathise—because, in my own early days of learning about Dr Deming’s
work, I was similarly mystified about this. So, to try to help, I have written some additional material just for
you which I’ll point you toward at the end of this section. But for now, as implied above, there is another
paradox to introduce, and I will deal with this second one right here in the main text.

Let me emphasise even further Deming’s concentration on the importance of control charts. He was clear
that it is generally desirable for *everyone* in an organisation to be able to use control charts and to communicate with them. And, even if that turns out not to be sensible or practical, he was also clear that, the
more senior someone is in their organisation, the more essential such ability becomes. As far as he was
concerned, the most important control charts should be right there on the Chief Executive’s desk.

Yet in his seminars Deming didn’t teach the details of how to construct control charts! Paradox No. 2.

Yes, he would usually draw up a chart once the results were obtained in his famous Red Beads Experiment
which we shall describe and work with tomorrow on Day 2 of our course. But how? He would just write
down a simple formula, insert some numbers that had been recorded during the experiment, and carry out
a little arithmetic. But there was nothing about where the simple formula came from nor the fact that, with
the large majority of processes, that same formula wouldn’t even be appropriate!

So let’s shed some light on this second paradox. This is a quote from my brief discussion on the sixth of
Deming’s famous 14 Points for Management: “Institute training” that you will see on Day 4 page 26 *[[WB 66]](workbook.html#sec-page66)*:

>In Deming’s terminology, the purpose of ‘training’ is the acquisition of specific skills for specific
tasks. Training is thus narrowly defined and finite in scale and scope. In contrast, as we shall see
in Point 13, ‘education’ is the opposite: very non-specific, very broadly defined, and essentially infinite in scale and scope.

Thus, in Deming’s terms, teaching and learning details about how to *construct* control charts is *training*.
Teaching and learning how to interpret the charts once they are drawn, and then how to proceed intelligently on the basis of such interpretation, is *education*. His seminars were most definitely *education*; what
he regarded as the simpler matter of training could be left to someone else. A couple of sentences from his diary entry for Monday 10 July 1950, regarding some lectures to the Japan Medical Association, will con-
firm the point:

><span class=deming_quote>Professor Masuyama and assistants will teach the statistical control of quality in the afternoon.
I shall teach during the forenoon the theory of a system, and cooperation.</span>

:::: {.columns}

::: {.column width="85%"}
So he regarded getting into some training, even on something as vital as the construction of control charts,
would be inappropriate during his teaching—it would interrupt the flow of what was really important: education. But in one respect his approach was dangerous. <span class=deming_quote>“Professor Masuyama and assistants”</span> might
have known what they were talking about but, sadly, many trainers who try to teach people about control
charts do not. Please be warned that much of what is currently taught and practised concerning control
charts is *not* wholly consistent with Shewhart’s teaching. For example, control charts are often regarded as
primarily relevant to manufacturing processes and requiring data that satisfy certain statistical conditions,
in particular that they are “normally distributed”. As previously, don’t worry if you do not know what that
means, for the fact is that Shewhart’s teaching was *not* dependent on such restrictions (though some statisticians seem to wish—or even believe—that it was). Quite the opposite. Indeed, Dr Shewhart was speci-
fic about the need to have statistical techniques which really work in practical situations rather than just
depend on what he referred to as “a fine ancestry of highbrow statistical theorems”! (This quotation comes
from page 18 of Shewhart’s 1931 book which, early this afternoon, you will see described in glowing terms
by Dr Deming.) A good way of expressing this is that Shewhart saw the need for statistical techniques that
are suitable for our data rather than the need for data that are suitable for our statistical techniques. Think
about it! Makes good sense to me. And that is precisely what Dr Shewhart produced.

:::

::: {.column width="15%"}
<div style="margin-top: 400px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(10, 00)
```
</div>

:::

::::

# Control charts in this course

So what does all this imply about my approach to control charts in this course? As in Dr Deming’s seminars, you will not have to construct any control charts if you’d prefer not to. However, you will see some
control charts on Days 2 and 3, and so I shall include some relevant description there. But there is already
plenty for you to do on both of those days, and so to insist that you also engage in any substantial training
on control charts would be too much of a good thing! On the other hand, the educational aspect that I’ve
already mentioned *is* absolutely vital. Repeating what I said above, that is to learn how to *interpret* the
charts once they are drawn, and then how to proceed intelligently on the basis of that interpretation.

On the other hand, you may of course *wish* to learn a little more or even a lot more about control charts. In
that case I’ll mention three possibilities which will enable you to go “beyond the basics”, in addition to Don
Wheeler’s book mentioned on page 5. The first is a small number of “Technical Aids” that appear in the
material for Days 2 and 3. These will enable you to construct the control charts yourself rather than just
use charts which other people have drawn up. The second possibility is an article which is freely available
on the internet. Some time ago I was encouraged by a freelance writer named Mitch Beedie to write an
introductory article on control charts. Mitch had found control charts extremely useful in his main areas of
interest: environmental studies and energy conservation. But he said that he had found introductory literature on the topic quite difficult to understand until he had come across some of mine! He titled the article
“Understanding Variation” and added the subtitle “The Springboard for Process Improvement”. I shall
therefore shall refer to it in this course material as the “Springboard” article. You can reach it on the UK
Deming Alliance’s website via the following link:

<https://demingalliance.org/resources/articles/understanding-variation-the-springboard-for-process-improvement>

And lastly, if you are *really* keen to have a comprehensive introduction to control charts—going well beyond
what is necessary for this course—there is the fairly substantial section of “Optional Extras” described at
the top of [page 8](index.qmd#para-page8) in the “Welcome” section. This optional material contains a variety of topics, all more or
less connected with control charting. But I must emphasise that both this extra material and the Springboard article really are “optional extras”—take them or leave them as you prefer. However, if you decide to
take them, please don’t count them as being within the recommended 12-day timeframe!

So how can I guide you through these various possibilities? I’m going to ask you to choose which of four
“Stats-levels” you feel is most suitable for you. Stats-level 0 refers to those who would prefer to avoid as
much as possible of anything like even simple arithmetic or drawing very basic graphs. Stats-level 1 is for
those who do not mind simple arithmetic and drawing basic graphs but who are not keen to go any further
than that. Stats-level 2 applies to those who would like to delve into something about the ideas underlying
the construction of control charts, but without getting in very deep. And Stats-level 3 is for those who are
already looking forward to tackling my Optional Extras section!

So which Stats-level are you?

After you’ve answered that question, here’s my guidance for you:

**Stats-level 0**: Don’t even bother with the few Technical Aids on Days 2 and 3. But you *will* be asked to
look at some pictures!

**Stats-level 1**: Work with the Technical Aids when they come, and have a go at constructing some of the
charts yourself.

**Stats-level 2**: As Stats-level 1 but also browse through the Springboard article when you feel like it and
take time to go through it more thoroughly after Day 3 in order to consolidate your learning.

**Stats-level 3**: As Stats-level 2 but take a look at what’s available in the Optional Extras section whenever
you like (out of hours!). Remember that this extra material is only there for those who might be genuinely
interested in some of it. Ignoring all of it will not harm your understanding of the rest of this course,
although some of it might help to increase your confidence about dealing with control charts when the time
comes that you need to. A sensible plan might therefore be to postpone the Optional Extras section until
the main 12-day course is completed—unless you just can’t resist looking at some of it earlier than that!

Having worked with some thousands of delegates at my seminars over the years, I am particularly conscious of the need to treat those on Stats-level 0 very gently! So there might be a case for going even further out in that direction and introducing a Stats-level 00! If you think that might apply to you then I should
also raise the possibility of you even skipping some of the main text in the morning of Day 3 (but nothing
else!)—I’ll guide you when you get there. Now, there are pros and cons in thinking of doing that, so please
bear with me while I discuss them.

As regards how knowledgeable somebody may become on control charts, it seems to me that there are
three stages. The first stage is the *only* absolutely vital one for this course. This is where you get to understand *why* control charts are important and what it is that they are able to communicate to their users.
Notice the “*why*” there rather than “*how*”. Getting as far as the “how” is the second stage. The third stage
goes even further and really is concerned with a further “why”: why is the “how” what it is?!

So the truth is that the first of the three stages is strictly all you need to know in order to be able to understand Dr Deming’s *theory* of management, which is primarily what this course is about. A later section of
this Overture takes a careful look at words like “theory”, so I won’t expand on that right now except to
make one crucial point. This is that, at least in Deming’s case, *theory* has a definite purpose—it is to guide
better *practice*. So, in Deming’s case, theory should be respected by *practical* people (possibly unlike
some other theories!). I am assuming that *you* are interested in better *practice* as well as learning the theory, so this will surely mean that, sooner or later, you *will* want to move onto the second stage: the “how”.

*How* can a control chart tell you what you need to know? That’s where you encounter the snag about skipping parts of the main text in the morning of Day 3. For, whereas they may be skipped if you were only
interested in theory, you *will* need to work through them when the time comes that you want to effectively
turn some of the theory into practice—the “better practice” that I just mentioned.

On the other hand, the third stage is *truly* optional as far as this course is concerned. Quite a lot of the
“how” is largely common sense, but some of it does depend on more “academic” information, such as a
few of the details needed to construct control charts. Faced with the choice, most people are content to
go by what the “experts” tell them is the sensible way, especially if there is some mathematics involved in
the whys and wherefores. So it is the second stage, the “how”, that is important to the *user* of control
charts—how to interpret what the control chart is telling you—rather than the fine detail of *why* the control
chart is constructed in the way that it is. If you want to watch television, do you really need to know why
the television—or even the remote control—works? Most people are content to read the instruction book
or, at least, the quick-start guide and do what they say. Some of that probably *is* common sense—but not
all of it. If the television doesn’t work for some reason, do you want to learn enough so that you can figure
out *why* it’s not working? No: you call in the engineer and trust that whoever responds to your call understands such things. That engineer has been educated and trained up to the third stage, whereas it is quite
sufficient for you and the large majority to go no further than the second stage.

So, in summary, unless you *really* want to limit yourself to Stats-level 00, do take at least a quick look at all
of the main text (except possibly for the Technical Aids). Even if you do not intend to start working with
control charts just yet, it would be a good idea for you to get some preliminary ideas about how to interpret
them. Then, when the time comes, you could come back to the practical guidance on interpreting control
charts (almost all of which is in the morning of Day 3) and study that more carefully then. However, if you
*are* verging on Stats-level 00, be content to just skim through that material for now: I don’t want you to get
so fed up with Day 3 that you feel discouraged and thus never move on to Day 4! I repeat: there won’t be
*any* involvement with control chart calculations etc throughout all the rest of this course after Day 3.

If you are following my timing guidance, the difference between Stats-level 0 (including 00, of course) and
the rest creates a small problem: on both Days 2 and 3, those on Stats-level 0 will have less to do than
everyone else. So alternative schedules will be provided on just those two days. On both days, the initial
Contents page will relate only to Stats-level 0 and will then be followed by a similar but different page for
Stats-levels 1–3. During the main text, the little clock icons for Stats-levels 1–3 will be on the right-hand
side as you are already seeing today, while the clock icons for Stats-level 0 will be on the left-hand side.

----------------------------------------------------------------------------

That finishes this section; so now, if you wish, take a little time out to read my discussion in the Appendix
on the former of the two paradoxes (which I described on page 7): I hope it will shed some light on the
mystery. But, again: that discussion is *only* intended for those who are puzzled by the paradox. If instead
you are simply content (and relieved!) by what the paradox has told you then please don’t bother with this
extra reading—just carry straight on. Also, the discussion on the first paradox is likely to be of greatest
interest to those who are familiar with the basics of “traditional” Statistics—for it includes one of Deming’s
most powerful expressions of the *irrelevance* of several of those basics when studying process data.

:::: {.columns}

::: {.column width="85%"}
So if you would now like to read about the first paradox on Appendix pages 2–6 and are also following my
guidance on timings at all closely then please “stop the clock” whilst you are reading those pages.
:::

::: {.column width="15%"}
<div style="margin-top: 40px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(10, 10)
```
</div>
:::

::::
