---
title: "THE ATTRACTION OF DEMING’S WORK"
execute:
  echo: false
---
  
Let me tell you now what first really began to attract me seriously to Deming’s work. It wasn’t the statistical
ideas (as you might have thought, because of my background as a statistician). It was the fact that his
teaching was so *human*! I saw and heard so much genuine, warm respect and empathy for humanity, individually and collectively. It was from him that I first heard ideas like <span class=deming_quote>“a company’s most important asset is
its people”</span>. But *he* really meant it (unlike some others from whom I’ve heard similar words). I heard scorn
of management whose first reaction when anything goes wrong is to look around for whom to blame, the
scapegoat. I heard passionate criticism of unemployment—or rather of management, and indeed government, that brings it about, be it accidental or intentional. It wasn’t that he was thinking about the cost of
supporting the unemployed. In contrast, it was the cost of the waste of human potential and self-respect
that concerned him. For confirmation, one could open his book *Out of the Crisis* at the very first page of its
*Preface*—there’s no need to even get into the main text! Here’s a brief extract[^a] from that page (in which I
wish the first few words were really true):

><span class=deming_quote>“It is no longer socially acceptable to dump employees on to the heap of unemployed. Loss of
market, and resulting unemployment, are not foreordained. They are not inevitable. They are manmade.”</span>

What else? Something that particularly appealed to me (I guess maybe because of my mathematical background) was the fact that his approach is eminently logical, scientific—in the best sense of those words.
I discovered (although not immediately) that everything he came up with was very carefully and logically
thought through, with plenty of supporting evidence. And I liked that—I don’t like taking things on faith, or
on trust—I never have: I want reasons. Call me a “Doubting Thomas” if you must.

And what else? His philosophy of management is wholly *system*- or *process*-orientated. It sees things
(including, very much, things to do with people) as full of interconnections, interdependences, interrelationships—as opposed to being largely isolated from each other. When I eventually realised that that was the
nature of Deming’s work, it rang very true to me. Consider e.g. what I just said about looking for a scapegoat. A natural consequence of his thinking (and Shewhart’s) is that, when something goes wrong, it’s far
more often the case that that’s due to the *system* within which people are working and living, rather than
being attributable to the people themselves. What obviously follows is that it’s far more often the case that
it’s the *system* that needs things doing to it, *not* the people. The Major Activity which concludes this first
day of the course is directly related to this aspect of Deming’s teaching.

I wonder if that surprises you. Don’t worry. This is still rather new thinking for many people. And, almost
by definition, new thinking is bound to be surprising.

I hope that this course will provide for you a good start to breaking through familiar old thinking, and to
catching more than a glimpse of what seems to me to be a very refreshing breakthrough from the depressing, demoralising, bureaucratic, judgmental, blame-ridden, dog-eat-dog, *inhuman* nature of what manifests
itself as “quality” in the view of much of modern management, and indeed modern government.

:::: {.columns}

::: {.column width="85%"}
So it is no accident that “government” features in the full title of Dr Deming’s final book: *The New Economics for Industry, Government, Education*. How, for example, can industry survive and succeed if it is hampered by poor government and education systems which do not understand the needs of the future?
:::

::: {.column width="15%"}
<div class="mt-sm">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(11, 45)
```
</div>
:::

::::

<div class="neave_note" role="note" aria-label="Author's note">Here is an example of a “request for action” that is exceedingly open-ended. You could content yourself
by writing some notes on the first few thoughts that come into your mind. Or it could be the subject of a
number of PhD theses! Since I’m calling this an “Activity” (not even a “*Major* Activity”!) I am indicating
that my preference lies toward the former of those two extremes!</div>

:::: {.columns}

::: {.column width="85%"}
<div class="thought" role="note" aria-label="Thought">
## ACTIVITY 1-d

Considering the breadth indicated in the title of Dr Deming’s final book: *The New Economics for Industry, Government, Education*, it will be worthwhile for you to spend a little time thinking about the ways
that the Government and Education systems in your country help or hinder Industry, and the quality of
what it produces and provides. How might they be improved? And, again referring to that title, how
might some “old” views of Economics also be a hindrance, and what might “New” views take into
account?

```{ojs}
viewof activity_1d = Inputs.textarea({label: "How do Government and Education systems help or hinder Industry?", placeholder: "Type your comments here.", rows: 30})
```

<div class=activity_afterthought>In such an Activity you are not expected to be able to set the world to rights! Its purpose is to trigger
some thoughts that might help you to make some useful connections as the course progresses.</div>
</div>
:::

::: {.column width="15%"}
<div class="spacer-3xl">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(12, 00)
```
</div>
:::

::::

:::: {.columns}

::: {.column width="85%"}
<div class="thought" role="note" aria-label="Thought">
## PAUSE FOR THOUGHT 1–e

<div class="neave_note" role="note" aria-label="Author's note">This is today’s final Pause for Thought and it goes to the opposite extreme compared with the previous Activity. Here I’m just asking you to consider briefly something very simple in preparation for discussion tomorrow. No calculations, no deep thought—just quickly jot down a couple of opinions for
future reference.</div>

Suppose you are the manager of a telephone call centre. A performance-monitoring scheme is in operation. For each member of staff who is being checked in any particular week, 50 of her callers are
phoned back and asked whether or not they were satisfied with how their enquiries had been dealt
with. The average satisfaction rate is around 80%, i.e. on average 40 of the callers turn out to have
been satisfied—or, looking on the negative side, on average 10 of the callers were dissatisfied.

At the end of each week you hold a “progress meeting” to comment on the results. Let’s look on the
grim side to start with! How much worse than the average of 10 dissatisfied callers would the result
need to be for you to feel justified in at least cautioning the member of staff concerned: “Watch it! Take
more care! This isn’t good enough!”? *Any* result worse than average, i.e. 11 or more dissatisfied customers? Actually, it’s not unknown for a person to be criticised if she gets just the 10 dissatisfied customers: after all, using a not-uncommon phrase, that’s “only average”.

<div class="neave_note" role="note" aria-label="Author's note">Please note that I am not indicating my approval of this or any other such scheme! To criticise people
simply because, according to some set of results, they appear below average or even “only average”
is, I suggest, a rather dubious judgment mechanism. After all, that means you’re bound to be criticising around half of them even if they are *all* performing brilliantly!</div>

Naturally, we would not expect everyone to get *exactly* 10 dissatisfied callers and 40 satisfied callers
every week (something to do with understanding variation!). Sometimes the result of the count will be
better than average, sometimes worse. But ! *how much* better or worse would it be reasonable for us to
expect “just by chance”?

So, to repeat the above question, how much worse than the average of 10 dissatisfied callers would the
result need to be for you to feel justified in at least cautioning the member of staff concerned? 12 dissatisfied callers? 13 dissatisfied callers? 14? What performance level do you feel, as the manager,
would warrant your attention? I am not asking for any logical argument here: just write down your “gut
feel”.

```{ojs}
viewof activity_1e_1 = Inputs.number([10, 50], {step: 1, label: "Dissatisfied callers", placeholder: 10})
```

On the bright side, how much better than the average of 10 dissatisfied callers would a result need to
be for you to feel justified in congratulating the member of staff: “Well done! Nice work! Keep it up!”?
9 dissatisfied callers? 8 dissatisfied callers? 7 dissatisfied callers? Fewer?

```{ojs}
viewof activity_1e_2 = Inputs.number([10, 0], {step: 1, label: "Dissatisfied callers", placeholder: 10})
```

</div>
:::

::: {.column width="15%"}
<div class="spacer-4xl">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(12, 10)
```
</div>
:::

::::

<div class="neave_note" role="note" aria-label="Author's note">
Let me briefly comment on the matter of averages. For in the case of people who cannot understand
averages, we shall have to work very hard to help them to understand variation!

Careless thought about the concept of “average” can have various peculiar consequences, either serious
or humorous depending upon circumstances. I relate a few near the end of *DemDim* Chapter 10: “Failures with Figures”. For example, Dr Deming was once told of the occasion when the Trades Union Con-
gress in Britain voted on the motion that no wage in the country should be below average. Apparently it
failed by just three votes. Deming wryly commented: <span class=deming_quote>“How I wish it had succeeded!”</span>. The TUC may have
been following the example of the Australian Minister of Labour who was quoted as having said in 1973
that “We look forward to the day when everybody will receive more than the average wage”. Even Jim
McIngvale, who will tell the story of his company Gallery Furniture here on Day 6, confesses when he fell
into the same trap: “Of the 80 salespeople we had, 10 to 15 turned over [*fired or quit*] every month—and
management wondered why half were still below average” ([Day 6 page 10](../day-06/02-gallery-furniture.qmd#sec-page10)).

Sadly, in recent years, I have heard British politicians making what seem to be similar judgments not just
about individuals but about hospitals and schools and branches of the police. The typical argument is
that those which are below average should have something done to them to bring them above average.
The other half are above average already, so why aren’t the rest? Think about it!
</div>

```{ojs download_all}
import { createDownloadButton } from "../../../assets/scripts/functions.js"

createDownloadButton([viewof activity_1d, viewof activity_1e_1, viewof activity_1e_2], "activities_1d_1e1_1e2.txt")
```

<!--
FOOTNOTES
-->

[^a]: This and all other quotations from *Out of the Crisis* have been reproduced with the approval of the
W Edwards Deming Institute and MIT Press. See <https://mitpress.mit.edu/contributors/w-edwards-deming>.