---
title: "THE DEADLY DISEASES"
execute:
  echo: false
---
Now it's time to take a break to watch the video *Management's Five Deadly Diseases* (see the link on page 1). Delegates used to appreciate my description of this video as featuring a rather elderly "Angry Young Man"! I think you'll soon see why.

<div class="return_callout">Watch *Management's Five Deadly Diseases* at [www.youtube.com/watch?v=ehMuuHGN0Y](https://www.youtube.com/watch?v=ehMuuHGN0Y) (note: the penultimate character is a zero; the sixth character from the end is a capital letter I).</div>

---

<div class="workbook_callout">The "Five Deadly Diseases" (pages 18–27) are also on Workbook pages 86–95.</div>

---

## Disease 1. Lack of constancy

<div class="info_callout">**The crippling disease is lack of constancy of purpose by planning to provide product and service in the future that will help man to live better materially, and which will have a market, and provide jobs.**</div>

In fact, Dr Deming began this statement with "The crippling disease in America…". I have omitted the two extra words for the obvious reason that this Disease is far more widespread.

We already know that "constancy of purpose" heads both the list of 14 Points and the list of Deadly Diseases, and how Dr Deming has thus focused upon its crucial importance—that's if his description: "the crippling disease" doesn't already have sufficient impact!

It is relevant here to point out the error in a phrase glibly trotted out by many speakers: "If you always do what you've always done, you'll always get what you've always got". *Wrong!* The world is ever-changing. Even the same products and services need to be produced more efficiently and cost-effectively: in other words, the *processes* involved still need to be improved—for, if you don't do it, others will. Secondly, of course, those same products and services may sooner or later have no market, because better products and services will have become available—and if not from you then from elsewhere. The market for even the best horse-drawn carriages soon plummeted once people were able to switch to cars. So it should be no surprise that Dr Deming's wording here is focused on *innovation*.

*(See Appendix page 28.)*

### Obsession With Quality

<div class="workbook_callout">Disease 1 (pages 18–19) is also on Workbook pages 86–87.</div>

```{ojs}
viewof d1_owq = Inputs.textarea({
  label: "Disease 1 (Lack of constancy) — Obsession With Quality",
  placeholder: "Your notes here.",
  rows: 5
})
```

### All One Team

```{ojs}
viewof d1_aot = Inputs.textarea({
  label: "Disease 1 (Lack of constancy) — All One Team",
  placeholder: "Your notes here.",
  rows: 5
})
```

### Scientific Approach

:::: {.columns}

::: {.column width="85%"}
```{ojs}
viewof d1_sa = Inputs.textarea({
  label: "Disease 1 (Lack of constancy) — Scientific Approach",
  placeholder: "Your notes here.",
  rows: 5
})
```
:::

::: {.column width="15%"}
<div style="margin-top: 20px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(12, 0)
```
</div>
:::

::::

---

## Disease 2. Short-termism

<div class="info_callout">**Short-term thinking defeats constancy of purpose to stay in business with long-term growth.**</div>

My choice of Dr Deming's words here are few but surely "hit the nail on the head".

My friend Dr Jackie Graham, based in Australia, points out how government cycles (i.e. the times between General Elections) can seriously obstruct long-term projects: "The big joke in Australia is the fast train from Melbourne to Sydney. Been talked about for more than 30 years. Every election time it gets rolled out, dusted off, and then after the election gets put back on ice. No government can build it in one term and so it does not get done." Jackie told me that one can drive a car between the two cities (without speeding) in considerably less time than the train currently takes! The latter barely averages 50 miles (80 kilometres) per hour. No wonder she observed: "the busiest air corridor in the world is between Melbourne and Sydney".

With even greater focus than the words I have quoted above, on *Out of the Crisis* page 84[99] Dr Deming expresses this Deadly Disease as "Emphasis on short-term profits". Short-term profits (like other figures discussed in Point 11 on page 10) can always be made to look better. In his first paragraph on this topic (*Out of the Crisis* page 85[99]), Dr Deming describes how to do it:

"Ship everything on hand, regardless of quality: mark it shipped, and show it all as accounts receivable. Defer till next quarter, so far as possible, orders for material and equipment. Cut down on research, education, training."

This will be so familiar to many readers that I need say no more.

*(See Appendix page 28.)*

### Obsession With Quality

<div class="workbook_callout">Disease 2 (pages 20–21) is also on Workbook pages 88–89.</div>

```{ojs}
viewof d2_owq = Inputs.textarea({
  label: "Disease 2 (Short-termism) — Obsession With Quality",
  placeholder: "Your notes here.",
  rows: 5
})
```

### All One Team

```{ojs}
viewof d2_aot = Inputs.textarea({
  label: "Disease 2 (Short-termism) — All One Team",
  placeholder: "Your notes here.",
  rows: 5
})
```

### Scientific Approach

:::: {.columns}

::: {.column width="85%"}
```{ojs}
viewof d2_sa = Inputs.textarea({
  label: "Disease 2 (Short-termism) — Scientific Approach",
  placeholder: "Your notes here.",
  rows: 5
})
```
:::

::: {.column width="15%"}
<div style="margin-top: 20px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(12, 0)
```
</div>
:::

::::

---

## Disease 3. Appraisal of performance

<div class="info_callout">**The effects of performance appraisal (personal review system, merit rating, evaluation of performance, annual review, system of reward, pay for performance, etc) are devastating.**</div>

As you can see from *DemDim* pages 50–51, Dr Deming's opinion of performance appraisal was unambiguously damning. But the term "performance appraisal" means different things to different people. So what exactly was Deming so heavily against, and why?

Basically, it was that judging individual performance with the implication that it is primarily the individual's *fault* if there are any supposed shortcomings. From Deming's viewpoint, one big reason is immediately and patently obvious: yet again, the vast majority of performance, behaviour, results, quality, productivity, etc comes from the system, not the individual. In particular, if appraisal means that a person will be rated or ranked—and, worse still, if this results in some kind of reward or punishment, be it financial and/or other—then that is what Deming was talking about and saw as so harmful.

As a personal observation, I can honestly say that I have never heard anybody who is involved with performance appraisal talking about it with any enthusiasm. From what people have so often told me, my impressions are these. For those to be appraised it is a matter of apprehension at least, even for those who usually escape pretty lightly from it—they might not be so lucky next time. For the appraisers it is an irritating chore. For all it is an unpleasant interruption to their "proper job".

Many organisations protest that this does not reflect *their* appraisal system—that is there to help the individual, for coaching, for seeing how management can assist the person to do a better job. If that is really the case, fine. But, if so, why only once a year? And also, if so, why call it "performance appraisal"? Consult your dictionary for its definitions of "appraise" and "appraisal".

It may be wise at this stage for me to remind you of my warning at the bottom of Appendix page 23 (see also my "Obsession With Quality" comment on Day 4 page 20 about the danger of merely stopping doing something undesirable and thus leaving a vacuum. In the context of performance appraisal, what is in my opinion the quite brilliant Chapter 9: "Performance Without Appraisal" in Peter Scholtes's *The Leader's Handbook* provides tremendously good guidance.

*(See Appendix page 29.)*

### Obsession With Quality

<div class="workbook_callout">Disease 3 (pages 22–23) is also on Workbook pages 90–91.</div>

```{ojs}
viewof d3_owq = Inputs.textarea({
  label: "Disease 3 (Appraisal of performance) — Obsession With Quality",
  placeholder: "Your notes here.",
  rows: 5
})
```

### All One Team

```{ojs}
viewof d3_aot = Inputs.textarea({
  label: "Disease 3 (Appraisal of performance) — All One Team",
  placeholder: "Your notes here.",
  rows: 5
})
```

### Scientific Approach

:::: {.columns}

::: {.column width="85%"}
```{ojs}
viewof d3_sa = Inputs.textarea({
  label: "Disease 3 (Appraisal of performance) — Scientific Approach",
  placeholder: "Your notes here.",
  rows: 5
})
```
:::

::: {.column width="15%"}
<div style="margin-top: 20px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(12, 0)
```
</div>
:::

::::

---

## Disease 4. Management job-hopping

<div class="info_callout">**Mobility of management causes instability, results in decisions being made by people who do not know the business** *[i.e. this business]* **and thus blindly use experience gained elsewhere which may be totally irrelevant.**</div>

This fourth Deadly Disease needs some clarification in order to avoid misinterpretation. Dr Deming is not suggesting that one should necessarily be stuck in the same place in the same organisation for life! But, as we are viewing an organisation as a system, broad knowledge across the organisation as a whole as well as substantial knowledge of part of it is *necessary*. How does the organisation work? How does what happens in one part of it affect what happens in other parts? And how is what happens in one part *affected by* what happens in other parts?

Moving around the organisation in order to acquire that broader understanding is clearly good. So what is the fourth Deadly Disease? What Dr Deming is warning us against here is the culture of flitting from one location to another, one job to another, one company to another, not for better understanding but for attempted self-advancement. In some environments you can be regarded as a "stick-in-the-mud" if you don't move on within a few months from wherever you are now. This encourages development of ability at short-term impact without concern for long-term effects (for you'll no longer be there); it encourages superficiality rather than depth; it encourages those actions without understanding. As so often the case, Dr Deming summed it up superbly in few words (*Out of the Crisis* page 103[121]):

"Mobility from one company to another creates prima donnas for quick results."

*(See Appendix page 29.)*

### Obsession With Quality

<div class="workbook_callout">Disease 4 (pages 24–25) is also on Workbook pages 92–93.</div>

```{ojs}
viewof d4_owq = Inputs.textarea({
  label: "Disease 4 (Management job-hopping) — Obsession With Quality",
  placeholder: "Your notes here.",
  rows: 5
})
```

### All One Team

```{ojs}
viewof d4_aot = Inputs.textarea({
  label: "Disease 4 (Management job-hopping) — All One Team",
  placeholder: "Your notes here.",
  rows: 5
})
```

### Scientific Approach

:::: {.columns}

::: {.column width="85%"}
```{ojs}
viewof d4_sa = Inputs.textarea({
  label: "Disease 4 (Management job-hopping) — Scientific Approach",
  placeholder: "Your notes here.",
  rows: 5
})
```
:::

::: {.column width="15%"}
<div style="margin-top: 20px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(12, 0)
```
</div>
:::

::::

---

## Disease 5. Use of only visible figures

<div class="info_callout">**One can not be successful on visible figures alone. Of course, visible figures are important: the bank account must be watched, and employees and vendors must be paid. But he who would run his company on visible figures alone will in time have neither company nor figures.**</div>

A familiar joke is "In God we trust; all others, bring data". Apart from religious objections some may have to this humour, it is dangerous in its implication that "data" are all that is needed for any purpose. By "data" is meant numerical information. Yet we have already become very familiar with Lloyd Nelson's claim that the most important figures for management of any organisation are "unknown and unknowable"—and now we have Deming drawing attention to the limitations of "visible figures".

This fifth Deadly Disease can come as a shock even to those who are warming to Deming's teachings—especially if they are statisticians or accountants! So what are these "most important" but invisible figures? You'll recall working on this in the third part of Activity 4-a (Day 4 page 5). Here is a summary of those mentioned by Dr Deming in his discussions on this Disease (*Out of the Crisis* pages 103–105[121–123] and elsewhere); they include…

"… measures of improvement in quality and productivity from continual improvement of processes, from elimination of work standards *[see the important note at the top of the next page]*, from better training, better supervision; from a team composed of the chosen supplier, the buyer, engineering design, sales, customer, working on a new component or redesign of an existing component; measures of loss from the annual rating on performance, and from inhibitors to pride of workmanship; measurements of the multiplying effect on sales that comes from a happy customer, and the opposite effect from an unhappy customer. Where are the figures?"

In conclusion, and back to the main issue (from *Out of the Crisis* pages 104–105 [page 123]):

"He that expects to quantify in dollars the gains that will accrue to a company year by year for a programme for improvement of quality by principles expounded in this book will suffer delusion. He should know before he starts that he will be able to quantify only a trivial part of the gain."

*(See Appendix page 30.)*

### Obsession With Quality

<div class="workbook_callout">Disease 5 (pages 26–27) is also on Workbook pages 94–95.</div>

```{ojs}
viewof d5_owq = Inputs.textarea({
  label: "Disease 5 (Use of only visible figures) — Obsession With Quality",
  placeholder: "Your notes here.",
  rows: 5
})
```

<div class="neave_note">**NB** There may be a potential danger in using the summary on the previous page without clarification: it is a possible misunderstanding of the term "work standards". But we only have to refer back to Point 11 (page 10): "work standards that prescribe quotas for the workforce and numerical goals for people in management" to be clear about Deming's focus with those words. That focus is, of course, in no way related to e.g. *standardising* work, procedures, operations, etc once they have been found advantageous. Apart from anything else, this clearly reduces variation. Some refer to these more appropriate practices by alternative terms such as "standard work" which helps to avoid possible ambiguity.</div>

### All One Team

```{ojs}
viewof d5_aot = Inputs.textarea({
  label: "Disease 5 (Use of only visible figures) — All One Team",
  placeholder: "Your notes here.",
  rows: 5
})
```

### Scientific Approach

:::: {.columns}

::: {.column width="85%"}
```{ojs}
viewof d5_sa = Inputs.textarea({
  label: "Disease 5 (Use of only visible figures) — Scientific Approach",
  placeholder: "Your notes here.",
  rows: 5
})
```
:::

::: {.column width="15%"}
<div style="margin-top: 20px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(12, 0)
```
</div>
:::

::::

---

That's it! Congratulations on completing this major project. You may now relax and look forward to a considerably shorter and less challenging Day 6! However, please see the note overleaf.

---

```{ojs download_all}
// Download button element
download_button = html`<button class="btn btn-primary" type="button">Download Your Notes</button>`
```

```{ojs download_trigger}
//| output: false

import { downloadNotes } from "../../../assets/scripts/functions.js"

download_button.onclick = () => {
  downloadNotes([
    viewof d1_owq, viewof d1_aot, viewof d1_sa,
    viewof d2_owq, viewof d2_aot, viewof d2_sa,
    viewof d3_owq, viewof d3_aot, viewof d3_sa,
    viewof d4_owq, viewof d4_aot, viewof d4_sa,
    viewof d5_owq, viewof d5_aot, viewof d5_sa
  ], "deadly_diseases.txt");
};
```

---

## "Out-of-hours" note

You may recall my saying in the Overture that I would, from time to time, ask you to focus attention on "your organisation" in some of this course's activities. The purpose is, of course, to help make your learning "live" in your own situation.

Day 6 is the first time that this happens—and it happens in quite a big way. However, it may of course be that you personally are not currently involved with any organisation, either as an employee or indeed as a manager—i.e. that there isn't a "your organisation" in this context. If you are studying this course as a member of a group then your discussions could perhaps centre around the organisation to which at least one of your fellow-students belongs. Failing that, and especially if you are studying alone, this will become the time to involve your friend—please refer back to the "Activities, Major Activities, and Pauses for Thought" section in the Overture: see Day 1 page 15.

The material for Day 6 includes a substantial case study on pages 7–16. Day 6's Major Activity then follows on pages 18–20. After you've read the first few paragraphs of the Major Activity, I'll suggest that you read through the case study again, but this time highlighting and making notes on items in the case study that will be helpful to you in the Major Activity—you'll probably be able to find quite a lot of such items. To help in that preparation I recommend that you print out in advance a single-sided "working copy" of Day 6 pages 7–16, and then staple or bind together those pages so that you can make notes on the blank sides opposite to the relevant items.

One further considerable help with this Major Activity would be, if possible, for you to print out an enlargement of the table on Day 6 page 19 *[WB 101]*. This is a table in which you will need to summarise the information you've just recorded alongside the case study, and so you would definitely benefit from having the extra space available on such an enlargement. Further guidance will be found in the middle of Day 6 page 18 *[WB 100]*.

**Approvals, Acknowledgments and Information**

<sup>a</sup> (page 14) This and all other quotations from Volume 21 of *The Deming Library* have been included with the approval of Clare Crawford-Mason.
