---
title: "Funnel-Rule Simulation Helpers — Developer Demo"
format:
  html:
    embed-resources: true
    toc: false
execute:
  echo: false
  warning: false
---

::: {.callout-note}
This is a **developer-facing fidelity check**, not a learner-facing page.
It is not linked from `_quarto.yml` and is intentionally scratch — the
consumer PR that swaps `assets/images/optional-extras/rule-1-short.png`
and `rule-1-long.png` for R-generated charts will delete this file.

Tracked under [issue #374](https://github.com/lddurbin/twelve_days_to_deming/issues/374).
:::

```{r setup, include=FALSE}
# The project-level .Rprofile sources R/setup.R at R session startup
# (from the repo-root working directory), so the four funnel_rule_*_sim()
# helpers and the chart wrapper are already on the search path here.
# This chunk is a no-op other than asserting that fact.
stopifnot(exists("funnel_rule_1_sim"),
          exists("funnel_simulation_chart_plot"))
```

## 2×2 panel: Rules 1–4 at default seeds

Each panel shows 50 drops of Neave's Funnel Experiment under one rule, with
MR-based control limits computed from the first 15 drops and extended across
the remaining 35. Target = 30, drop noise sd = 5 — the Optional Extras Part A
scale.

```{r demo-panel, fig.width=10, fig.height=8}
library(patchwork)

n_total    <- 50
n_baseline <- 15
sd_drop    <- 5
target     <- 30

mk <- function(values, title, y_limits, y_breaks) {
  funnel_simulation_chart_plot(
    values,
    n_baseline     = n_baseline,
    y_limits       = y_limits,
    y_breaks       = y_breaks,
    y_minor_breaks = seq(min(y_limits), max(y_limits), by = 1),
    line_width     = 1.5
  ) +
    ggplot2::ggtitle(title) +
    ggplot2::theme(
      plot.title    = ggplot2::element_text(hjust = 0.5, face = "bold",
                                            size = 13, colour = CHART_FG),
      axis.text.y   = ggplot2::element_text(size = 9, colour = CHART_FG),
      axis.text.x   = ggplot2::element_text(size = 9, colour = CHART_FG)
    )
}

r1 <- mk(funnel_rule_1_sim(n_total, sd = sd_drop, target = target),
        "Rule 1 — funnel fixed at target",
        y_limits = c(10, 50), y_breaks = seq(10, 50, by = 10))
r2 <- mk(funnel_rule_2_sim(n_total, sd = sd_drop, target = target),
        "Rule 2 — compensate by last deviation",
        y_limits = c(10, 50), y_breaks = seq(10, 50, by = 10))
r3 <- mk(funnel_rule_3_sim(n_total, sd = sd_drop, target = target),
        "Rule 3 — mirror across target (explodes)",
        y_limits = c(-30, 90), y_breaks = seq(-30, 90, by = 30))
r4 <- mk(funnel_rule_4_sim(n_total, sd = sd_drop, target = target),
        "Rule 4 — funnel follows the marble (drift)",
        y_limits = c(-10, 70), y_breaks = seq(-10, 70, by = 20))

(r1 | r2) / (r3 | r4)
```

## What to look for

- **Rule 1** — points cluster around the central line (target = 30), staying
  comfortably inside the MR-based control limits across all 50 drops. Should
  look like one of Neave's six stable processes on Day 3 page 19.
- **Rule 2** — pronounced central zig-zag; the MR-inflated control limits sit
  *wider* than they should and the run-chart stays well inside them, the
  "hugging the Central Line" effect Neave describes on Optional Extras
  Part A page 10.
- **Rule 3** — amplitude grows visibly across the 50 drops; the run-chart
  eventually pierces the early-baseline limits.
- **Rule 4** — the marble drifts away from target. Short-term moving ranges
  are small (so the limits hug the early-baseline values), but later
  drops wander well outside them — the "wandering" effect Neave describes
  on Optional Extras Part A page 13.

## Reproducibility

Each call uses each helper's documented default seed (Rule 1 = 7, Rule 2 = 2,
Rule 3 = 1, Rule 4 = 42). Re-render this qmd and the figure is bit-for-bit
identical.
