---
title: "INTRODUCTION TO THE FUNNEL EXPERIMENT"
execute:
  echo: false
---

```{r}
# R setup for all activities (edit as needed)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
source(here::here("R/functions/main-functions.R"))
```

# Activity Heading

Introduction to the Funnel Experiment

This is a major activity that will help you understand variation.

Timing: 15:00

Pause for thought: Consider what you know about variation.

Your turn: Write down your thoughts about the experiment.

Download your notes when complete.


:::: {.columns}
::: {.column width="85%"}
Content with timing indicator.
:::
::: {.column width="15%"}
<div style="margin-top: 10px">
```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_clock(15, 00)
```
</div>
:::
::::


```{ojs}
viewof activity_id = Inputs.textarea({placeholder: "Type your comments here.", rows: 10})
```

<button class="btn btn-primary" type="button" data-bs-toggle="collapse" data-bs-target="#collapse_feedback" aria-expanded="false" aria-controls="collapse_feedback">
Show Commentary
</button>
<div class="collapse" id="collapse_feedback">
Activity feedback or commentary goes here.
</div>

<div class=activity_afterthought>Optional afterthought or tip.</div>

```{ojs download_all}
// Download button element
// (You can add more viewof variables to the array as needed)
download_button = html`<button class="btn btn-primary" type="button">Download Your Notes</button>`
```

```{ojs download_trigger}
//| output: false
import { downloadNotes } from "/assets/scripts/functions.js"
download_button.onclick = () => {
  downloadNotes([viewof activity_id], "activity_notes.txt");
};
``` 